GSE158422: Package Development

### 2023-11-21 figs/GSE158422_pkgdev.Rmd
a walkthrough for infinium limma users
Prefer UCSC annotations

tic()
anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
myann <- data.frame(anno[,c("GencodeCompV12_NAME","GencodeCompV12_Group","Islands_Name","Relation_to_Island")])
gp <- myann[,"GencodeCompV12_NAME",drop=FALSE]
gp2 <- strsplit(gp$GencodeCompV12_NAME,";")
names(gp2) <- rownames(gp)
gp2 <- lapply(gp2,unique)
gt <- stack(gp2)
colnames(gt) <- c("gene","probe")
dim(gt)
str(gt)
toc() #9.0s

@@

tic()
anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
myann <- data.frame(anno[,c("GencodeCompV12_NAME","GencodeCompV12_Group","Islands_Name","Relation_to_Island")])
gp <- myann[,"GencodeCompV12_NAME",drop=FALSE]
gp2 <- strsplit(gp$GencodeCompV12_NAME,";")
names(gp2) <- rownames(gp)
gp2 <- lapply(gp2,unique)
gt <- stack(gp2)
colnames(gt) <- c("gene","probe")
dim(gt)
str(gt)
toc() #2.44, 2.41, 2.41s

@@

mitch_report(res=mres,outfile="GSE158422_pkgdev_mitchreport.html",overwrite=TRUE)
mitch_plot2(res=mres,outfile="GSE158422_pkgdev_mitchcharts.pdf")

@@

Now do the same but filter for promoters.
Using the GencodeCompV12_NAME field of the annotation and using TSS probes only.

@@

head(myann)
table(anno$Regulatory_Feature_Group)
table(unlist(strsplit(myann$GencodeCompV12_Group,";")))

# promoter prep
tic()
keep=c("TSS200","TSS1500")
annos <- strsplit(myann$GencodeCompV12_Group,";")
genes <- strsplit(myann$GencodeCompV12_NAME,";")
keeps <- lapply(1:nrow(myann), function(i) {
  unique(genes[[i]][annos[[i]] %in% keep])
})
names(keeps) <- rownames(myann)
#gt <- as.data.frame((unlist(gp2)))
gt2 <- stack(gp2)
colnames(gt2) <- c("gene","probe")
toc() #20.37, 19.56, 21.23

>>>>

tic()
anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
myann <- data.frame(anno[,c("UCSC_RefGene_Name","GencodeCompV12_Group","Islands_Name","Relation_to_Island")])
gp <- myann[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
gp2 <- lapply(gp2,unique)
gt <- stack(gp2)
colnames(gt) <- c("gene","probe")
dim(gt)
str(gt)
toc() #9.0s

@@

tic()
anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
myann <- data.frame(anno[,c("UCSC_RefGene_Name","GencodeCompV12_Group","Islands_Name","Relation_to_Island")])
gp <- myann[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
gp2 <- lapply(gp2,unique)
gt <- stack(gp2)
colnames(gt) <- c("gene","probe")
dim(gt)
str(gt)
toc() #2.44, 2.41, 2.41s

@@

mitch_report(res=mres,outfile="GSE158422_pkgdev_mitchreport.html",overwrite=TRUE)
mitch_plots(res=mres,outfile="GSE158422_pkgdev_mitchcharts.pdf")

@@

Now do the same but filter for promoters.
Using the UCSC_RefGene_Name field of the annotation and using TSS probes only.

@@

head(myann)
table(anno$UCSC_RefGene_Group)
table(unlist(strsplit(myann$GencodeCompV12_Group,";")))

# promoter prep
tic()
keep=c("TSS200","TSS1500")
annos <- strsplit(myann$GencodeCompV12_Group,";")
genes <- strsplit(myann$UCSC_RefGene_Name,";")
keeps <- lapply(1:nrow(myann), function(i) {
  unique(genes[[i]][annos[[i]] %in% keep])
})
names(keeps) <- rownames(myann)
#gt <- as.data.frame((unlist(gp2)))
gt2 <- stack(gp2)
colnames(gt2) <- c("gene","probe")
toc() #20.37, 19.56, 21.23
